---
title: "Demonstration of Law of Large Numbers and Central Limit Theorem using exponential distribution."
author: "TG Shannon"
date: "December 15, 2015"
output: 
  pdf_document: 
    latex_engine: xelatex
---

# Overview

Law of Large Numbers tells us the mean of a sample of size n converges to mu, the population mean, as n becomes increasingly large with a probability of 100%.

The Central Limit Theorem tells us that the mean of a large number of iterations of independent random variables, each with a well defined mean and variance, will be approximately normally distributed, regardless of the underlying distribution.

The report will demonstrate the Central Limit Theorem using the exponential distribution. We will use a lambda parameter = 0.2 for the simulation. Therefore, the theoritical mean and standard deviation is 1/0.2 = 5.

# Simulations

```{r}
lambda <- 0.2
N <- 1000
n <- 40
population <- rexp(N, lambda)
hist(population, main = "Distribution of Simulated Population")
```

The mean of the simulated population is `r mean(population)` and the standard deviation is `r sd(population)` compared to the theoritical values of `r 1/lambda`. The simulated population is a good representation of the theoritical population, and is clearly not normally distributed.

```{r}
samples <- matrix(rexp(N*n,lambda), N)
sample_vars <- apply(samples, 1, var)
sample_means <- apply(samples, 1, mean)
h<-hist(sample_means, breaks=10, density=10, col="gray", xlab="Sample mean", main="Distribution of Sample Means")
xfit<-seq(min(sample_means),max(sample_means),length=n)
yfit<-dnorm(xfit,mean=mean(sample_means),sd=sd(sample_means))
yfit <- yfit*diff(h$mids[1:2])*length(sample_means)
lines(xfit, yfit, col="blue", lwd=2)
```

The mean of the sample means is `r mean(sample_means)` compared with the theorical population mean of `r 1/lambda`, and the predicted standard error of the mean is `r 1/lambda/sqrt(n)` compared with the calculated value of `r sd(sample_means)`. The 95% confidence interval for the mean of the sample means distribution is 
`r mean(sample_means) + c(-1,1)*qnorm(.975) * 2*sd(sample_means)/sqrt(length(sample_means)) `,
which contains the theoritical population mean of `r 1/lambda`.

The distribution of sample means is shown to be normally distributed with sample mean equal to the population mean, and sample standard distribution equal to the population standard error for a sample size of `r n`.